Converting Neural Networks to Gaussian Processes for Sequential Decision-Making Under Uncertainty

### 2023-04-21 paper/main.tex
added sentance to abstract
Minor
RL figure placeholder

\documentclass{article}

@@

@@

\begin{abstract}
% Building systems that adapt to new information and express uncertainty is essential to deploy methods in real-world applications e.g. healthcare, autonomous vehicles. While gradient-based training is symbiotic to modern deep learning, it is rigid regarding adding further information, usually involving retraining from scratch. Converting a trained neural network to its dual parameterization we can view the parameters of the model as data dependent terms instead of the weights. Furthermore, we can then build a sparse model to form a variational dual sparse Gaussian process. The resultant sparse GP has two obvious advantages: scalable principled uncertainty estimation and the ability adapt to new data without gradient based retraining. Our proposed method can be deployed in any sequential decision-making process, where new information is received and retraining from scratch is not available. We show its effectiveness in model-based reinforcement learning using updated information for more efficient exploration.
%
  Building learning-based systems that can adapt to new data and express uncertainty is essential for their deployment in real-world applications, e.g. healthcare and robotics. However, modern deep learning methods cannot adapt to new data because they rely on gradient-based optimisation, which requires retraining with the addition of new data. In contrast, nonparametric approaches -- such as Gaussian processes -- can adapt to new data. However, they are limited to low-dimensional and small data sets. We present a method that enables any trained neural network (NN) to be converted to a sparse variational Gaussian process (SVGP). Importantly, our method 1) enables fast adaptation to new data, 2) provides well-calibrated uncertainty estimates, and 3) scales to high-dimensional and large data sets. Essentially, our method reaps the benefits of nonparametric models without their well-known issues. We demonstrate the effectiveness of our model's fast adaptation by showing that it can improve sample-efficiency in model-based reinforcement learning and we demonstrate our state-of-the-art NN uncertainty quantification with experiments using UCI data sets.

@@

\section{Experiments}

@@

\section{Conclusion} \label{sec:conclusion}

\section*{Broader Impact}

@@

@@

\section{Appendix}

@@

\begin{algorithm}[!t]
\caption{Model-based RL}\label{alg-mbrl}
\begin{algorithmic}[1]
  \Require Start state $\state_{0}$, initial data set $\dataset_{0}$, dynamics posterior $p(\transitionFn \mid \dataset_{0})$, policy $\policy_{0}$
\For{$i  \in \{1, 2, \ldots, \text{num episodes} \}$}
    \State Reset the system to $\state_{0}$ and reset trajectory buffers $\bm\tau_{t} = \emptyset \ \forall t$
    \For{$t  \in \{1, 2, \ldots, \text{num steps} \}$}
      % \State Collect  $\tau_{0:t} = \tau_{0:t-1} \cup (\state_{j}, \action_{j}, \state_{j+1}, r_{j+1})$
      \State Use \cref{eq-greedy}/\cref{eq-posterior-sampling}/\cref{eq-ucrl} to collect data $\bm\tau_{t} = \bm\tau_{t-1} \cup (\state_{t}, \policy_{i}(\state_{t}), \transitionFn(\state_{t}, \policy_{i}(\state_{t})), r_{t+1})$
      % \State Execute policy $\policy_{i}(\state_{t})$ in environment and update trajectory $\tau_{i+1} = \{\state_{j}, \action_{j}, \state_{j+1}, r_{j+1}) \}_{j=0}^{t}$
    \EndFor
    \State Update data set $\dataset_{0:i} = \dataset_{0:i-1} \cup \tau$
    \State Train dynamics $p(\transitionFn \mid \dataset_{0:i}) \leftarrow \text{update\_dynamics}(\dataset_{0:i}, p(\transitionFn \mid \dataset_{0:i-1}))$
    % \State Train dynamics $p(\transitionFn \mid \dataset_{0:i+1})$ using $\dataset_{0:i+1}$
    % \State Improve policy $\pi_{i+1}$ using $p(\transitionFn \mid \dataset_{0:i+1})$ and/or $\dataset_{0:i+1}$
    \State Improve policy $\pi_{i+1} \leftarrow \text{update\_policy}(p\left(\transitionFn \mid \dataset_{0:i}), \dataset_{0:i} \right)$
    %\State Improve policy $\pi_{i+1}$ using $p(\transitionFn \mid \dataset_{0:i+1})$ and/or $\dataset_{0:i+1}$
\EndFor
\end{algorithmic}
\end{algorithm}

>>>>

% !TeX spellcheck = en_GB
\documentclass{article}

@@

% Redefine paragraph to be tighter
\renewcommand{\paragraph}[1]{{\bf #1}~~}

% Latin
\usepackage{xspace}
\newcommand{\eg}{\textit{e.g.\@}\xspace}
\newcommand{\ie}{\textit{i.e.\@}\xspace}
\newcommand{\cf}{\textit{cf.\@}\xspace}
\newcommand{\etc}{\textit{etc.\@}\xspace}
\newcommand{\etal}{\textit{et~al.\@}\xspace}

% Tikz
\usepackage{tikz}
\usepackage{pgfplots}
\usetikzlibrary{patterns}
\usetikzlibrary{decorations,backgrounds,arrows.meta,calc}

% Appendix/supplement title
\newcommand{\nipstitle}[1]{{%
    % rules for title box at top and bottom
    \def\toptitlebar{\hrule height4pt \vskip .25in \vskip -\parskip} 
    \def\bottomtitlebar{\vskip .29in \vskip -\parskip \hrule height1pt \vskip .09in} 
    \phantomsection\hsize\textwidth\linewidth\hsize%
    \vskip 0.1in%
    \toptitlebar%
    \begin{minipage}{\textwidth}%
        \centering{\LARGE\bf #1\par}%
    \end{minipage}%
    \bottomtitlebar%
    \addcontentsline{toc}{section}{#1}%
}}

@@

\begin{abstract}
% Building systems that adapt to new information and express uncertainty is essential to deploy methods in real-world applications e.g. healthcare, autonomous vehicles. While gradient-based training is symbiotic to modern deep learning, it is rigid regarding adding further information, usually involving retraining from scratch. Converting a trained neural network to its dual parameterization we can view the parameters of the model as data dependent terms instead of the weights. Furthermore, we can then build a sparse model to form a variational dual sparse Gaussian process. The resultant sparse GP has two obvious advantages: scalable principled uncertainty estimation and the ability adapt to new data without gradient based retraining. Our proposed method can be deployed in any sequential decision-making process, where new information is received and retraining from scratch is not available. We show its effectiveness in model-based reinforcement learning using updated information for more efficient exploration.
%
The world is a dynamic environment that requires humans and animals to constantly adapt.
  Building learning-based systems that can adapt to new data and express uncertainty is essential for their deployment in real-world applications, e.g. healthcare and robotics. However, modern deep learning methods cannot adapt to new data because they rely on gradient-based optimisation, which requires retraining with the addition of new data. In contrast, nonparametric approaches -- such as Gaussian processes -- can adapt to new data. However, they are limited to low-dimensional and small data sets. We present a method that enables any trained neural network (NN) to be converted to a sparse variational Gaussian process (SVGP). Importantly, our method 1) enables fast adaptation to new data, 2) provides well-calibrated uncertainty estimates, and 3) scales to high-dimensional and large data sets. Essentially, our method reaps the benefits of nonparametric models without their well-known issues. We demonstrate the effectiveness of our model's fast adaptation by showing that it can improve sample-efficiency in model-based reinforcement learning and we demonstrate our state-of-the-art NN uncertainty quantification with experiments using UCI data sets.

@@

\section{Experiments}
\label{sec:experiments}

\todo{Overview of experiments}

\subsection{Benchmarks on UCI to for supervised learning}

\subsection{Supervised learning on image data sets}

\subsection{Updating the representation during continual learning}

\subsection{Fast updates in reinforcement learning}

\begin{figure}[!t]
  \centering
  \begin{subfigure}{.2\textwidth}
  \resizebox{\textwidth}{!}{%
  \begin{tikzpicture}[inner sep=0,outer sep=0]

    % Draw decorated 'ground'
    \draw[postaction={draw, decorate, decoration={border, angle=-45,
					amplitude=1cm, segment length=.5cm}}] (-3,-1.5) -- (12,-1.5);

    % The cart
    \draw[draw=black,fill=black!50,draw=black,line width=3pt,rounded corners=1mm] (0,0) rectangle (9cm,3cm);
    \node[fill=black,circle,minimum size=.5cm] (dot) at (4.5cm,3cm) {};

    % Wheels
    \node[fill=white,draw=black,line width=3pt,circle,minimum size=2cm,,fill=black!50] at (2cm,-.5cm) {};
    \node[fill=white,draw=black,line width=3pt,circle,minimum size=2cm,fill=black!50] at (7cm,-.5cm) {};

    % The arm
    \node[anchor=north,minimum width=1cm,minimum height=14cm,draw=black,rotate=20,rounded corners=5mm,yshift=7mm,xshift=-.3mm,fill=white,draw=black,line width=3pt] at (dot) {};
    \node[fill=black,circle,minimum size=.5cm] at (dot) {};

    % Markings
    \draw[loosely dashed,line width=1pt] (4.5,6) -- (4.5,-10);

    % Arrow
    \draw[->,black,line width=3pt,-{Latex[length=7mm,width=7mm]}] (1,5) --node[above,outer sep=6pt]{\large $x$} (4.5,5);

    \def\centerarc[#1](#2)(#3:#4:#5)% Syntax: [draw options] (center) (initial angle:final angle:radius)
    { \draw[#1] ($(#2)+({#5*cos(#3)},{#5*sin(#3)})$) arc (#3:#4:#5); }

    % Draw arc
    \centerarc[black,line width=3pt](dot)(270:290:11)
    \node at (6.5,-9) {$\theta$};

    \node[white,minimum size=1cm] at (4.5,-13) {};

  \end{tikzpicture}}
  \end{subfigure}
  \hfill
  \begin{subfigure}{.78\textwidth}
    \tikz\node[minimum width=\textwidth,minimum height=.35\textwidth,fill=black!10,rounded corners=2pt]{RL trjectories or something interesting?};
  \end{subfigure}
  \hfill
  \caption{Add caption.}
  \label{fig:rl}
\end{figure}

@@

\section{Discussion and Conclusion}
\label{sec:conclusion}

%\section*{Broader Impact}

@@

\nipstitle{
    {\Large Supplementary Material:} \\
    PAPER TITLE GOES HERE}
\pagestyle{empty}

@@

\section{Method Details}

@@

\begin{algorithm}[!b]
\caption{Model-based RL}\label{alg-mbrl}
\begin{algorithmic}[1]
  \Require Start state $\state_{0}$, initial data set $\dataset_{0}$, dynamics posterior $p(\transitionFn \mid \dataset_{0})$, policy $\policy_{0}$
\For{$i  \in \{1, 2, \ldots, \text{num episodes} \}$}
    \State Reset the system to $\state_{0}$ and reset trajectory buffers $\bm\tau_{t} = \emptyset \ \forall t$
    \For{$t  \in \{1, 2, \ldots, \text{num steps} \}$}
      % \State Collect  $\tau_{0:t} = \tau_{0:t-1} \cup (\state_{j}, \action_{j}, \state_{j+1}, r_{j+1})$
      \State Use \cref{eq-greedy}/\cref{eq-posterior-sampling}/\cref{eq-ucrl} to collect data $\bm\tau_{t} = \bm\tau_{t-1} \cup (\state_{t}, \policy_{i}(\state_{t}), \transitionFn(\state_{t}, \policy_{i}(\state_{t})), r_{t+1})$
      % \State Execute policy $\policy_{i}(\state_{t})$ in environment and update trajectory $\tau_{i+1} = \{\state_{j}, \action_{j}, \state_{j+1}, r_{j+1}) \}_{j=0}^{t}$
    \EndFor
    \State Update data set $\dataset_{0:i} = \dataset_{0:i-1} \cup \tau$
    \State Train dynamics $p(\transitionFn \mid \dataset_{0:i}) \leftarrow \text{update\_dynamics}(\dataset_{0:i}, p(\transitionFn \mid \dataset_{0:i-1}))$
    % \State Train dynamics $p(\transitionFn \mid \dataset_{0:i+1})$ using $\dataset_{0:i+1}$
    % \State Improve policy $\pi_{i+1}$ using $p(\transitionFn \mid \dataset_{0:i+1})$ and/or $\dataset_{0:i+1}$
    \State Improve policy $\pi_{i+1} \leftarrow \text{update\_policy}(p\left(\transitionFn \mid \dataset_{0:i}), \dataset_{0:i} \right)$
    %\State Improve policy $\pi_{i+1}$ using $p(\transitionFn \mid \dataset_{0:i+1})$ and/or $\dataset_{0:i+1}$
\EndFor
\end{algorithmic}
\end{algorithm}
